---
title: "Demonstration"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# library(sfreader)
```

# Overview
The purpose of the `sfreader` package is to read Summary Files from the
decennial census and write the contents to a relational database (such as
SQLite) which can be easily queried. Because Summary Files each have subtly
different formats, each requires its own reader to be written.

This is a personal project of the author, and is not endorsed in any way
by the U.S. Census Bureau. The package is also currently under heavy
construction, and should not be depended upon in any serius way.

# Installation
To install from Github, obtain the `devtools` package and run the following in R.

```{r, eval=FALSE}
devtools::install_github("andrewraim/sfreader")
```

# Example: Create a Database for 2010 Oklahoma SF2 files

First, browse to the repository for SF2 files:
<https://www.census.gov/data/datasets/2010/dec/summary-file-2.html>
Download the file `ok2010.sf2.zip` which contains data for Oklahoma.
Unzip the contents to a folder on your local computer, say
`C:/Users/raim0001/Documents/datasets/sf2_2010/Oklahoma/`.
Now we will initialize an `sf2_2010_reader` object and point it to
the files.

```{r, echo=TRUE, results='hide', message=FALSE, warning=FALSE}
library(sfreader)
reader = sf2_2010_reader$new("C:/Users/raim0001/Documents/datasets/sf2_2010/Oklahoma/")
```

Our destination will be a SQLite database which resides in a file `sf2_2010.sqlite`.
Let's create the database there. To avoid using too much space, let's just save 
PCT1 (Total Population) and PCT3 (Sex x Age).
```{r, echo=TRUE, results='hide', message=FALSE, warning=FALSE}
library(RSQLite)
sqlite_file = "sf2_2010.sqlite"
unlink(sqlite_file)
conn = dbConnect(RSQLite::SQLite(), sqlite_file)
reader$write_sqlite(conn, target_tables = c("PCT1", "PCT3"))
dbDisconnect(conn)
```

# Example: Using the Database
Now we should be able to connect to the database whenever we like and query
it. We don't necessarily need to do this through R - we could use any kind of client 
which can connect to the database. First, let's request the list of tables.
```{r}
conn = dbConnect(RSQLite::SQLite(), sqlite_file)
dbListTables(conn)
```

Here is a brief description of all the tables:  
- `Geo_XYZ`: Geo table for summary level `XYZ`. For example, `Geo_040` represents states.  
- `Iterations`: List of iterations (race classifications).  
- `SummaryLevels`: List of available summary levels.  
- `TableNames`: Names and descriptions of tables in this summary file.  
- `PCT1`: The table `PCT1` for all summary levels and all iterations.  
- `PCT3`: The table `PCT3` for all summary levels and all iterations.  

We can list fields in some of those tables.
```{r}
dbListFields(conn, "Geo_050")
dbListFields(conn, "PCT1")
```

Here are a few simple queries.
```{r}
q_out = dbGetQuery(conn, statement = "select NAME, LOGRECNO from Geo_040")
head(q_out)
q_out = dbGetQuery(conn, statement = "select `Iteration Code` from Iterations")
head(q_out)
```

And here are a few queries involving multiple tables.
```{r}
q_out = dbGetQuery(conn, statement = "
	select g.REGION, g.DIVISION, g.STATE, g.COUNTY, p.*
	from Geo_050 g, PCT1 p
	where g.FILEID == p.FILEID
	and g.LOGRECNO == p.LOGRECNO
")
head(q_out, 3)

q_out = dbGetQuery(conn, statement = "
	select g.REGION, g.DIVISION, g.STATE, g.COUNTY, p.CHARITER, p.LOGRECNO,
		p.PCT0010001, it.Iterations
	from Geo_050 g, PCT1 p, Iterations it
	where g.FILEID == p.FILEID
	and g.LOGRECNO == p.LOGRECNO
	and p.CHARITER == it.`Iteration Code`
")
head(q_out, 3)
```
